Hochschild cohomology of torus equivariant D-modules: New approach

### 2017-10-10 rewrite.tex
Fiber products, part 1

@@

\begin{Def}
    Let $\stack X$ be any QCA stack.
    Consider the Cartesian diagram
    \[
        \begin{tikzcd}
            \ls \stack X \arrow[d, "p_1"] \arrow[r, "p_2"] & \stack X \arrow[d, "\Delta"] \\
            \stack X \arrow[r, "\Delta"] & \stack X \times \stack X
        \end{tikzcd}
    \]
    We say that $\stack X$ has property \bc\ if the canonical morphism
    \[
        p_{1,!}p_2^! \to \Delta^!\Delta_!
    \]
    of endofunctors of $\catDModHol{\stack X}$ is an equivalence.
\end{Def}

@@

\begin{Def}
    Let $\stack X$ be a QCA stack and consider the cartesian square
    \[
        \begin{tikzcd}
            \stack G_{n+1} \arrow[r, "p_{s}"] \arrow[d, "p_{F+1}"] & \stack G_n \arrow[d, "p_F"] \\
            \stack G_{m+1} \arrow[r, "p_s"] & \stack G_m
        \end{tikzcd}
    \]
    where the horizontal maps are induced by a map $F\colon [n] \to [m]$ in $\catCoSimplicial$.
    We say that $\stack X$ has property \hbc\ if for each such $F$ the canonical morphism
    \[
        p_{F+1,!}p_s^! \to p_s^!p_{F,!}
    \]
    of functors $\catDModHol{\stack G_n} \to \catDModHol{\stack G_{m+1}}$ is an equivalence.
\end{Def}

\begin{Cor}\label{cor:properties_imply_monads}
    If $\stack X$ has properties \bc\ and \hbc, then there exists a canonical structure of monad on $p_{1,!}p₂^!$ and the morphism $p_{2,!}p₁^! → Δ^!Δ_!$ is an isomorphism of monads.
    In particular there is an isomorphism of algebras
    \[
        \HCoh\bigl(\catDMod{\stack X}\bigr)
        \cong
        \opalg{\GammadR\bigl(\stack X,\, p_{1,!} p₂^! k_{\stack X}\bigr)}.
    \]
\end{Cor}

@@

\begin{proof}
    We will show the statement for property \bc.
    The proof for \hbc\ is essentally the same.
    Let us first assume that $i$ is open.
    It suffices to show that $p_{\stack Y,1,!}p_{\stack Y,2}^!i^! \to \Delta_{\stack Y}^!\Delta_{\stack Y,!}i^!$ is an equivalence.
    Now, using the assumption on $\stack X$ and the fact that $i^! = i^*$ we get
    \begin{multline*}
        p_{\stack Y,1,!} p_{\stack Y,2}^!i^! =
        p_{\stack Y,1,!} (\ls i)^! p_{\stack X,2}^! \cong
        i^! p_{\stack X,1,!} p_{\stack X,2}^! \isoto \\
        i^! \Delta_{\stack X}^! \Delta_{\stack X,!} =
        \Delta_{\stack Y}^! (i\times i)^! \Delta_{\stack X,!} \cong
        \Delta_{\stack Y}^! \Delta_{\stack X,!} i^!.
    \end{multline*}
    Now let $i$ be a closed immersion.
    As above, it suffices to show that $i_!p_{\stack Y,1,!}p_{\stack Y,2}^!i^! \to i_!\Delta_{\stack Y}^!\Delta_{\stack Y,!}i^!$ is an equivalence.
    Now, using the assumption on $\stack X$ and the fact that $i_! = i_*$ we get
    \begin{multline*}
        i_! p_{\stack Y,1,!} p_{\stack Y,2}^!i^! =
        p_{\stack X,1,!} (\ls i)_! p_{\stack Y,2}^!i^! \cong
        p_{\stack X,1,!} p_{\stack X,2}^! i_!i^! \isoto \\
        \Delta_{\stack X}^! \Delta_{\stack X,!} i_!i^! \cong
        \Delta_{\stack Y}^! (i\times i)_! \Delta_{\stack X,!}i^! =
        i_! \Delta_{\stack Y}^! \Delta_{\stack X,!} i^!.
        \qedhere
    \end{multline*}
\end{proof}

@@

\begin{Lem}\label{lem:open_cover}
    If $\stack X$ has an cover by open substacks $\stack U_i$ with property \bc\ (resp.~with property \hbc), then $\stack X$ has property \bc\ (resp.~property \hbc).
\end{Lem}

@@

\begin{proof}
    The category $\catDMod{\stack X_1 \times \stack X_2}$ is generated by elements of the form $\sheaf F_1 \boxtimes \sheaf F_2$ with $\sheaf F_i \in \catDModHol{\stack X_i}$.
    For any schematic morphism $f\colon \stack Y \to \stack Z$ of QCA stacks the functor $f^!$ is continuous on $\catDMod{\stack Z}$ and $f_!$ is preservers colimits in $\catDModHol{\stack Y}$, since it has a right adjoint.
    Since the inclusion of a full subcategory reflects all colimits it is thus sufficient to check that the properties holds on elements of the form $\sheaf F_1 \boxtimes \sheaf F_2 \in \catDModHol{\stack X_1 \times \stack X_2}$.

@@

@@

\begin{proof}
    By Sumihiro's theorem \cite{Sumihiro:1974:EquivariantCompletions} and Lemma~\ref{lem:open_cover} we can assume that $\stack X$ is affine with a linear action of $T$.
    By Lemma~\ref{lem:lc_immersion} it further suffices to consider the case that $\stack X = \as{m}/T$ for some $n$.
    By Lemma~\ref{lem:product}, we can further decompose $\as{m}$ into $T$-eigenspaces.
    Each of those splits into copies of $\as{1}/T \cong \as{1}/\Gm \times BT'$.

    By Lemma~\ref{lem:smooth_or_proper} it remains to consider the stack $\as{1}/\Gm$, where $\Gm$ acts nontrivially.
    If the $\Gm$ action is contractive this case follows immediately from Proposition~\ref{prop:contractive} and Lemma~\ref{lem:smooth_or_proper}.
    Otherwise the inverse of the $\Gm$-action on $\as1$ is contractive and we can use the contraction principle (and hence Proposition~\ref{prop:contractive}) with the monoid $\as1 = \Gm \cup \{\infty\}$ instead of $\Gm \cup \{0\}$.
\end{proof}

>>>>

\newcommand\catK[2][]{\cat{K}_{#1}(#2)}
\newcommand\catKHol[1]{\catK[\mathrm{hol}]{#1}}

@@

In order for the isomorphism \eqref{eq:intro:naive} to hold, we only need know that the base change morphism is an isomorphism on $k_{\stack X}$.
However, it is often useful to pass to substacks, so that we need to prove the isomorphism of a slightly larger class of D-modules.
Thus we introduce the following subcategories.

\begin{Def}
  Let $\catK{\stack X}$ be the full subcategory of $\catDMod{\stack X}$ generated by $k_{\stack X}$ and let $\catKHol{\stack X}$ be its full subcategory consisting of holonomic objects.
\end{Def}

With this we can write down the two properties that we need to check in order to get a good

\begin{Def}
    Let $\stack X$ be any QCA stack.
    Consider the Cartesian diagram
    \[
        \begin{tikzcd}
            \ls \stack X \arrow[d, "p_1"] \arrow[r, "p_2"] & \stack X \arrow[d, "\Delta"] \\
            \stack X \arrow[r, "\Delta"] & \stack X \times \stack X
        \end{tikzcd}
    \]
    We say that $\stack X$ has property \bc\ if the canonical morphism
    \[
        p_{1,!}p_2^! \to \Delta^!\Delta_!
    \]
    of functors $\catKHol{\stack X} \to \catDModHol{\stack X}$ is an equivalence.
\end{Def}

@@

\begin{Def}
    Let $\stack X$ be a QCA stack and consider the cartesian square
    \[
        \begin{tikzcd}
            \stack G_{n+1} \arrow[r, "p_{s}"] \arrow[d, "p_{F+1}"] & \stack G_n \arrow[d, "p_F"] \\
            \stack G_{m+1} \arrow[r, "p_s"] & \stack G_m
        \end{tikzcd}
    \]
    where the horizontal maps are induced by a map $F\colon [n] \to [m]$ in $\catCoSimplicial$.
    We say that $\stack X$ has property \hbc\ if for each such $F$ the canonical morphism
    \[
        p_{F+1,!}p_s^! \to p_s^!p_{F,!}
    \]
    of functors $\catKHol{\stack G_n} \to \catDModHol{\stack G_{m+1}}$ is an equivalence.
\end{Def}

\begin{Cor}\label{cor:properties_imply_monads}
  \remark{This needs to be fixed as we don't have endofunctors anymore.}
    If $\stack X$ has properties \bc\ and \hbc, then there exists a canonical structure of monad on $p_{1,!}p₂^!$ and the morphism $p_{2,!}p₁^! → Δ^!Δ_!$ is an isomorphism of monads.
    In particular there is an isomorphism of algebras
    \[
        \HCoh\bigl(\catDMod{\stack X}\bigr)
        \cong
        \opalg{\GammadR\bigl(\stack X,\, p_{1,!} p₂^! k_{\stack X}\bigr)}.
    \]
\end{Cor}

@@

\begin{proof}
    We will show the statement for property \bc.
    The proof for \hbc\ is essentially the same.
    First we note that if $\sheaf F \in \catKHol{\stack Y}$, then $i_*\sheaf F \in \catKHol{\stack X}$ since
    \[
      \Hom_{\catDMod{\stack X}}(k_{\stack X}, i_*\sheaf F) \cong
      \Hom_{\catDMod{\stack Y}}(i^*k_{\stack X}, \sheaf F) \cong
      \Hom_{\catDMod{\stack Y}}(k_{\stack Y}, \sheaf F) \ncong 0.
    \]
    Thus it suffices to show that $p_{\stack Y,1,!}p_{\stack Y,2}^!i^! \to \Delta_{\stack Y}^!\Delta_{\stack Y,!}i^!$ is an isomorphism of $\catKHol{\stack X}$.
    We split $i$ into a closed immersion followed by an open one and prove the two cases separately.

    Let us first assume that $i$ is open.
    Now, using the assumption on $\stack X$ and the fact that $i^! = i^*$ we get
    \begin{multline*}
        p_{\stack Y,1,!} p_{\stack Y,2}^!i^! =
        p_{\stack Y,1,!} (\ls i)^! p_{\stack X,2}^! \cong
        i^! p_{\stack X,1,!} p_{\stack X,2}^! \isoto \\
        i^! \Delta_{\stack X}^! \Delta_{\stack X,!} =
        \Delta_{\stack Y}^! (i\times i)^! \Delta_{\stack X,!} \cong
        \Delta_{\stack Y}^! \Delta_{\stack X,!} i^!.
    \end{multline*}
    Now let $i$ be a closed immersion.
    As above, it suffices to show that $i_!p_{\stack Y,1,!}p_{\stack Y,2}^!i^! \to i_!\Delta_{\stack Y}^!\Delta_{\stack Y,!}i^!$ is an equivalence.
    Now, using the assumption on $\stack X$ and the fact that $i_! = i_*$ we get
    \begin{multline*}
        i_! p_{\stack Y,1,!} p_{\stack Y,2}^!i^! =
        p_{\stack X,1,!} (\ls i)_! p_{\stack Y,2}^!i^! \cong
        p_{\stack X,1,!} p_{\stack X,2}^! i_!i^! \isoto \\
        \Delta_{\stack X}^! \Delta_{\stack X,!} i_!i^! \cong
        \Delta_{\stack Y}^! (i\times i)_! \Delta_{\stack X,!}i^! =
        i_! \Delta_{\stack Y}^! \Delta_{\stack X,!} i^!.
        \qedhere
    \end{multline*}
\end{proof}

@@

\begin{Lem}\label{lem:open_cover}
    If $\stack X$ has an cover by open substacks $\stack U_i$ with property \bc\ (resp.~with property \hbc), then $\stack X$ has property \bc\ (resp.~property \hbc).
    \qed
\end{Lem}

@@

\begin{proof}
    The category $\catDMod{\stack X_1 \times \stack X_2}$ is generated by elements of the form $\sheaf F_1 \boxtimes \sheaf F_2$ with $\sheaf F_i \in \catDModHol{\stack X_i}$.
    For any schematic morphism $f\colon \stack Y \to \stack Z$ of QCA stacks the functor $f^!$ is continuous on $\catDMod{\stack Z}$ and $f_!$ is preservers colimits in $\catDModHol{\stack Y}$, since it has a right adjoint.
    Since the inclusion of a full subcategory reflects all colimits it is thus sufficient to check that the properties holds on elements of the form $\sheaf F_1 \boxtimes \sheaf F_2 \in \catKHol{\stack X_1 \times \stack X_2}$.

@@

We would like to have an extension of Lemma~\ref{lem:product} to fiber products.
Unfortunately, as the discussion on \cite[Section~1.2]{BenZviNadler:arXiv:CharacterTheoryOfAComplexGroup} shows, the category of D-modules on a fiber product $\stack X_1 \times_{\stack Y} \stack X_2$ is rather badly behaved unless $\stack Y$ is a scheme.
Thus we have to restrict ourselves to the following partial result which is sufficient for our applications.

\begin{Lem}\label{lem:product_quotient}
  Let $G$ be an affine algebraic group acting on schemes $X_1$ and $X_2$.
  Suppose the quotient stacks $\stack X_1 = X_1/G$ and $\stack X_2 = X_2/G$ both have property \bc\ (resp.~property \hbc).
  Then so does $(X_1 \times X_2)/G$.
\end{Lem}

\begin{proof}
  The stack $(X_1 \times X_2)/G$ can be rewritten as $\stack X_1 \times_{\cs{G}} \stack X_2$.
  Thus the results of \cite[Section~5.2]{BenZviNadler:arXiv:CharacterTheoryOfAComplexGroup} show that $!$-pullback exhibits the category $\catDMod{\stack X_1} \otimes_{\catDMod{\cs{G}}} \catDMod{\stack X_2}$ as a full subcategory of $\catDMod{\stack X_1 \times_{\cs{G}} \stack X_2}$ containing $\catK{\stack X_1 \times_{\cs{G}} \stack X_2}$.
  This category can be described as the limit of the cosimplicial category with cosimplices
  \[
    \catDMod{\stack X_1 \times \cs{G} \times \dotsb \times \cs{G} \times \stack X_2}.
  \]
  Thus it suffices to show the properties for the stacks $\stack X_1 \times \cs{G} \times \dotsb \times \cs{G} \times \stack X_2$, where they follow form Lemmas~\ref{lem:product} and~\ref{lem:smooth_or_proper}\ref{lem:smooth_or_proper:bg}.
\end{proof}

@@

\begin{proof}
  By Sumihiro's theorem \cite{Sumihiro:1974:EquivariantCompletions}\remark{exact reference?} and Lemma~\ref{lem:open_cover} we can assume that $\stack X$ is affine with a linear action of $T$.
  By Lemma~\ref{lem:lc_immersion} it further suffices to consider the case that $\stack X = \as{m}/T$ for some $n$.
  By Lemma~\ref{lem:product_quotient}, we can further decompose $\as{m}$ into $T$-eigenspaces.
  Each of those splits into copies of $\as{1}/T$, which in turn can be written as $\as{1}/\Gm \times BT'$ for some subtorus $T'$ of $T$.

  By Lemma~\ref{lem:smooth_or_proper}\ref{lem:smooth_or_proper:bg} it remains to consider the stack $\as{1}/\Gm$, where $\Gm$ acts nontrivially.
  If the $\Gm$ action is contractive this case follows immediately from Proposition~\ref{prop:contractive} and Lemma~\ref{lem:smooth_or_proper}.
  Otherwise the inverse of the $\Gm$-action on $\as1$ is contractive and we can use the contraction principle (and hence Proposition~\ref{prop:contractive}) with the monoid $\as1 = \Gm \cup \{\infty\}$ instead of $\Gm \cup \{0\}$.
\end{proof}
